Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-02-18 paper/rebuttal/rebuttal.tex

The flow $\hat{T_{ij}}$ is symmetric (\ie $\hat{T_{ij}}=\hat{T_{ji}}$) because {\it word2vec} neglects whether the context $j$ appears before or after the center $i$ in the trajectory---the mechanism of skip-gram.
If we swap $i$ and $j$ in Eq.~\eqref{eq:flow_w2v_ng1}, the numerator remains the same but the denominator can be different.
Therefore, to ensure $\hat{T}_{ij} = \hat{T}_{ji}$, the denominator $Z_i$ should be a constant.

Taken altogether, the {\it word2vec} model with the negative sampling predicts a flow in the same form as in the gravity model:
\begin{align}
	\hat{T_{ij}}=  C P(i) P(j)  \exp(\bm{v}_j \cdot \bm{v}_{i}), \label{eq:flow_w2v_ng0}
\end{align}
where $C$ is a positive constant.

In other words, \emph{word2vec with the negative sampling is a gravity model}, with the mass given by the location's frequency $P(i)$, and the distance measured by their dot similarities.
While the gravity model generates mobility flow from the given mass and locations, {\it word2vec} finds the mass and locations that best explain the given mobility flow.
Therefore, the embedding distance is inherently tied to mobility flow, and hence, has greater predictive power than the geographic distance.

A central assumption in our mathematical analysis is that the embedding vectors are optimum in terms of the {\it word2vec}'s objective function.
However, this premise is not practical because the optimization algorithm (\eg stochastic gradient descent algorithm) often finds sub-optimal embeddings.
This has two consequences.
First, the in- and out-vectors can be different.
Second, dot similarity calculated by {\it word2vec} is the distance metric tied to the gravity model, and hence, should capture the mobility flow well.
Our numerical simulations show that dot similarity calculated by {\it word2vec} is the second-best predictor of mobility flow among the ten embedding distances~(Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}).
However, cosine similarity predicted mobility flow better than dot similarity.
This result echos many applications of word embedding such as word analogy test~\autocite{levy2015improving}, in which cosine similarity outperforms dot similarity.
In predicting mobility flow, if the embedding vectors are optimal, cosine similarity should be no better than dot similarity because cosine similarity is a special case of dot similarity.
Therefore, the inferiority of dot similarity is attributed to the optimization process, and a careful tuning of the optimization algorithm would close the gap between our mathematical and numerical results.

### 2021-02-18 paper/current/main.tex
fix minot type: last sentence of math

A central assumption in our mathematical analysis is that the embedding vectors are optimum in terms of the {\it word2vec}'s objective function.
However, this premise is not practical because the optimization algorithm (\eg stochastic gradient descent algorithm) often finds sub-optimal embeddings.
This has two consequences.
First, the in- and out-vectors can be different.
Second, dot similarity calculated by {\it word2vec} is the distance metric tied to the gravity model, and hence, should capture the mobility flow well.
Our numerical simulations show that dot similarity calculated by {\it word2vec} is the second-best predictor of mobility flow among the ten embedding distances~(Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}).
However, cosine similarity predicted mobility flow better than dot similarity.
This result echos many applications of word embedding such as word analogy test~\autocite{levy2015improving}, in which cosine similarity outperforms dot similarity.
In predicting mobility flow, if the embedding vectors are optimal, cosine similarity should be no better than dot similarity because cosine similarity is a special case of dot similarity.
Therefore, the inferiority of dot similarity is attributed to the optimization process, and therefore, a careful tuning of the optimization algorithm would close the gap between our mathematical and numerical results.

>>>>

A central assumption in our mathematical analysis is that the embedding vectors are optimum in terms of the {\it word2vec}'s objective function.
However, this premise is not practical because the optimization algorithm (\eg stochastic gradient descent algorithm) often finds sub-optimal embeddings.
This has two consequences.
First, the in- and out-vectors can be different.
Second, dot similarity calculated by {\it word2vec} is the distance metric tied to the gravity model, and hence, should capture the mobility flow well.
Our numerical simulations show that dot similarity calculated by {\it word2vec} is the second-best predictor of mobility flow among the ten embedding distances~(Tables~\ref{supp:table:r2_table},~\ref{supp:table:rmse_exp_table},~\ref{supp:table:rmse_power_table}).
However, cosine similarity predicted mobility flow better than dot similarity.
This result echos many applications of word embedding such as word analogy test~\autocite{levy2015improving}, in which cosine similarity outperforms dot similarity.
In predicting mobility flow, if the embedding vectors are optimal, cosine similarity should be no better than dot similarity because cosine similarity is a special case of dot similarity.
Therefore, the inferiority of dot similarity is attributed to the optimization process, and a careful tuning of the optimization algorithm would close the gap between our mathematical and numerical results.
